### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@cbfd5cdbdcfbacffeec53d9f575e8be2dd9348a7 🚀
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@9853410081187b5af0c9187e2a3e047b0cb4c087 🚀

Poisot \emph{et al.} (2012) and Canard \emph{et al.} (2014) have
suggested an approach to \(\beta\)-diversity for ecological networks
which is based on the comparison of the number of shared and unique
links among species within a pair of networks. Their approach
differentiates this sharing of links between those established between
species occurring in both networks, and those established with at least
one unique species. This framework is expressed as the decomposition
\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that network
dissimilarity (\(\beta_{wn}\)) has a component that can be calculated
directly from the dissimilarity of interactions between shared species
(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)).
Presumably, the value of these components for a pair of networks can
generate insights about the mechanisms involved in dissimilarity.

@@

Yet, the precise meaning of \(\beta_{st}\), namely the importance of
species turnover in the overall dissimilarity, has been difficult to
capture, and a source of confusion for some practitioners. This is not
particularly surprising, as this component of the decomposition responds
to unique species introducing their unique interactions both between
themselves, and with species that are common to both networks
fig.~\ref{fig:conceptual}. For this reason, it is important to come up
with guidelines for the interpretation of this measure, and how to use
it to extract ecological insights.

@@

In the following sections, I present a series of calculations aimed at
expressing the values of \(\beta_{os}\), \(\beta_{wn}\), and therefore
\(\beta_{st}\) as a function of species sharing probability (as a proxy
for mechanisms generating turnover), and link rewiring probability (as a
proxy for mechanisms generating differences in interactions among shared
species). These calculations are done using \texttt{Symbolics.jl} (Gowda
\emph{et al.} 2021), and subsequently transformed in executable code for
\emph{Julia} (\textbf{Bezanson2017JulFre?}), used to produce the
figures.

@@

So as to simplify the notation of the following section, I will
introduce a series of new variables. Let \(C = |E_c|\) be the number of
links that are identical between networks (as a mnemonic, \(C\) stands
for ``common''); \(R = |E_{sn} \cup E_{sm}|\) be the number of links
that are not shared, but only involve shared species (\emph{i.e.} links
from \(\mathcal{M}\cup\mathcal{N}\) established between species from
\(\mathcal{M}\cap\mathcal{N}\); as a mnemonic, \(R\) stands for
``rewired''); and \(T = |E_{un} \cup E_{um}|\) the number of links that
are not shared, and involve at least one unique species (as a mnemonic,
\(T\) stands for ``turnover'').

There are two important points to note here. First, as mentionned
earlier, the number or proportion of species that are shared is not
involved in the calculation. Second, the connectance of either network
is not involved in the calculation. That all links counted in
\emph{e.g.} \(U\) come from \(\mathcal{M}\), or that they are evenly
distributed between \(\mathcal{M}\) and \(\mathcal{N}\), has no impact
on the result. This is a desirable property of the approach: whatever
quantitative value of the components of dissimilarity can be interpreted
in the light of the connectance and species turnover \emph{without} any
risk of circularity; indeed, I present a numerical experiment where
connectance varies independently later in this manuscript, reinforcing
this point.

@@

Using the \(\beta_t\) measure of dissimilarity, we can re-write (using
the notation with \(A\), \(S\), and \(U\))

@@

>>>>

Poisot \emph{et al.} (2012) and Canard \emph{et al.} (2014) have
suggested an approach to \(\beta\)-diversity for ecological networks
which is based on the comparison of the number of shared and unique
links among species within a pair of networks. Their approach
differentiates this sharing of links between those established between
species occurring in both networks, and those established with at least
one unique species. This framework is expressed as the decomposition
\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that network
dissimilarity (\(\beta_{wn}\)) has a component that can be calculated
directly from the dissimilarity of interactions between shared species
(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)). The
\(\beta_{st}\) component differs slightly from the others, in that it is
a quantification of the \emph{relative} rewiring to overall
dissimilarity, and not an \emph{absolute} measure of interaction
turnover. Presumably, the value of these components for a pair of
networks can generate insights about the mechanisms involved in
dissimilarity, when interpreted within the context of species turnover
and differences in network connectance.

@@

Yet, the precise meaning of \(\beta_{st}\), namely the importance of
species turnover in the overall dissimilarity, has been difficult to
capture, and a source of confusion for some practitioners. This is not
particularly surprising, as this component of the decomposition responds
to unique species introducing their unique interactions both between
themselves, and with species that are common to both networks
(fig.~\ref{fig:conceptual}). For this reason, it is important to come up
with guidelines for the interpretation of this measure, and how to use
it to extract ecological insights.

@@

In the following sections, I present a series of calculations aimed at
expressing the values of \(\beta_{os}\), \(\beta_{wn}\), and therefore
\(\beta_{st}\) as a function of species sharing probability (as a proxy
for mechanisms generating turnover), and link rewiring probability (as a
proxy for mechanisms generating differences in interactions among shared
species). These calculations are done using \texttt{Symbolics.jl} (Gowda
\emph{et al.} 2021), and subsequently transformed in executable code for
\emph{Julia} (Bezanson \emph{et al.} 2017), used to produce the figures.

@@

Note that this measure is written as \(\beta_t\) for consistency with
Koleff \emph{et al.} (2003). So as to simplify the notation of the
following section, I will introduce a series of new variables. Let
\(C = |E_c|\) be the number of links that are identical between networks
(as a mnemonic, \(C\) stands for ``common'');
\(R = |E_{sn} \cup E_{sm}|\) be the number of links that are not shared,
but only involve shared species (\emph{i.e.} links from
\(\mathcal{M}\cup\mathcal{N}\) established between species from
\(\mathcal{M}\cap\mathcal{N}\); as a mnemonic, \(R\) stands for
``rewired''); and \(T = |E_{un} \cup E_{um}|\) the number of links that
are not shared, and involve at least one unique species (as a mnemonic,
\(T\) stands for ``turnover'').

There are two important points to note here. First, as mentionned
earlier, the number or proportion of species that are shared is not
involved in the calculation. Second, the connectance of either network
is not involved in the calculation. That all links counted in
\emph{e.g.} \(T\) come from \(\mathcal{M}\), or that they are evenly
distributed between \(\mathcal{M}\) and \(\mathcal{N}\), has no impact
on the result. This is a desirable property of the approach: whatever
quantitative value of the components of dissimilarity can be interpreted
in the light of the connectance and species turnover \emph{without} any
risk of circularity; indeed, I present a numerical experiment where
connectance varies independently later in this manuscript, reinforcing
this point.

@@

Using the \(\beta_t\) measure of dissimilarity, we can re-write (using
the notation with \(R\), \(C\), and \(T\))

@@

\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
Bezanson, J., Edelman, A., Karpinski, S. \& Shah, V. (2017). Julia: A
Fresh Approach to Numerical Computing. \emph{SIAM Review}, 59, 65--98.
